---
title: "Activism Pilot Analysis"
author: "Shaye"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
    theme: flatly
---

# Set Up

```{r setup + set wd, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
options(scipen = 999, digits = 4)

suppressPackageStartupMessages({
  library(tidyverse)
  library(here)
  library(lmtest)
  library(sandwich)
  library(emmeans)
  library(lme4)
  library(lmerTest)
  library(sjPlot)
  library(ggeffects)
  library(ggthemes)
  library(ggpubr)
  library(patchwork)
  library(kableExtra)
  library(gtsummary)
  library(rempsyc)
  library(psych)
  library(interactions)
  library(effectsize)
  library(ggsignif)
  library(paletteer)
  library(report)
  library(modelsummary)
  library(corrplot)
  library(ggdist)
  library(performance)
  library(survival)
  
  # Econometric & simulation packages
  for (pkg in c("AER", "estimatr", "simr", "clubSandwich", "fixest", "fwildclusterboot")) {
    if (requireNamespace(pkg, quietly = TRUE)) {
      suppressWarnings(suppressPackageStartupMessages(library(pkg, character.only = TRUE)))
    }
  }
})

# Robust path resolver across VS Code, RStudio, and command-line execution
find_path <- function(...) {
  p1 <- here::here(...)
  if (file.exists(p1) || dir.exists(p1)) return(p1)
  
  p2 <- here::here("analysis", ...)
  if (file.exists(p2) || dir.exists(p2)) return(p2)
  
  p3 <- file.path(...)
  if (file.exists(p3) || dir.exists(p3)) return(p3)
  
  p4 <- file.path("..", ...)
  if (file.exists(p4) || dir.exists(p4)) return(p4)

  return(p1)
}

# Repo root -- tables/results and figures/results are git-tracked at the repo
# root and read by paper.tex via root-relative paths, regardless of where
# here::here() resolves (analysis/ under the .Rproj, or elsewhere on the CLI)
repo_root <- trimws(system("git rev-parse --show-toplevel", intern = TRUE))
paper_path <- function(...) file.path(repo_root, ...)

# Holm-adjusted p-values for every lm_robust model passed to modelsummary(), applied
# per model across all non-intercept terms (matches the paper's stated methodology:
# "we apply Holm-adjusted p-values throughout"). modelsummary's own estimate/SE/gof
# (AIC, BIC, RMSE, R2...) are left untouched -- only the p.value column is overridden,
# via modelsummary's tidy_custom S3 hook.
tidy_custom.lm_robust <- function(x, ...) {
  pp <- as.data.frame(parameters::model_parameters(x))
  is_intercept <- pp$Parameter == "(Intercept)"
  p_adj <- pp$p
  p_adj[!is_intercept] <- p.adjust(p_adj[!is_intercept], method = "holm")
  data.frame(term = pp$Parameter, p.value = p_adj)
}

# Same Holm-adjustment for the lmer (lmerTest) interaction models -- fixed effects
# only; random-effect variance rows have no p-value and are left for modelsummary's
# own native gof handling.
tidy_custom.lmerModLmerTest <- function(x, ...) {
  pp <- as.data.frame(parameters::model_parameters(x, effects = "fixed"))
  is_intercept <- pp$Parameter == "(Intercept)"
  p_adj <- pp$p
  p_adj[!is_intercept] <- p.adjust(p_adj[!is_intercept], method = "holm")
  data.frame(term = pp$Parameter, p.value = p_adj)
}

# Pretty term labels matching the paper's established style -- factor dummies as
# "condition [level]", lag variables with underscores read as spaces
REG_TERM_LABELS <- c(
  "conditionpetition"       = "condition [petition]",
  "conditiondemonstration"  = "condition [demonstration]",
  "conditionroadblock"      = "condition [roadblock]",
  "conditioncontrol"        = "condition [control]",
  "others_contribution_lag" = "others contribution lag",
  "NegDev_lag"               = "NegDev lag",
  "PosDev_lag"               = "PosDev lag",
  # interaction models (treatment_assigned/is_treatment_round, demonstration = reference)
  "treatment_assignedpetition"                   = "treatment [petition]",
  "treatment_assignedroadblock"                  = "treatment [roadblock]",
  "is_treatment_round"                            = "treatment round",
  "treatment_assignedpetition:is_treatment_round" = "treatment [petition] x treatment round",
  "treatment_assignedroadblock:is_treatment_round" = "treatment [roadblock] x treatment round"
)

# modelsummary's LaTeX output wraps the table in its own \begin{table}...\end{table};
# paper.tex already wraps each \input{} in its own table float (caption, label,
# resizebox), so strip that outer wrapper here to avoid nesting two table floats.
strip_table_wrapper <- function(path) {
  ln <- readLines(path)
  ln <- ln[!ln %in% c("\\begin{table}", "\\centering", "\\end{table}")]
  writeLines(ln, path)
}

# estimate/statistic/shape put p in its own column next to "Estimate (SE)", matching
# the paper's established table style, instead of modelsummary's default of stacking
# SE and p together underneath the estimate within a single per-model column.
save_reg_table <- function(models, path, coef_rename = REG_TERM_LABELS,
                            estimate = "{estimate} ({std.error})",
                            statistic = "p.value",
                            shape = term ~ model + statistic,
                            ...) {
  modelsummary::modelsummary(models, output = path, coef_rename = coef_rename,
                              estimate = estimate, statistic = statistic, shape = shape, ...)
  strip_table_wrapper(path)
}
```

# Load Processed Data

```{r load data}
# Load processed clean datasets (generated by 01_data_cleaning.Rmd)
load(find_path("data", "processed", "clean_dfs.RData"))
```

# Descriptive Statistics

```{r desc outcomes}
#outcomes by condition

df_long %>%
  filter(!is.na(relative_contribution)) %>%
  group_by(treatment_assigned,is_treatment_round) %>%
  rename(Condition = treatment_assigned,
         "Treatment Round" = is_treatment_round) %>%
  summarise(
    n                = n(),
    "Avg Contribution %" = mean(relative_contribution, na.rm = TRUE),
    "Avg Contribution % (SD)"   = sd(relative_contribution, na.rm = TRUE),
    "Avg # of Activists"    = mean(num_activists, na.rm = TRUE),
    "Avg Activism Rate"    = mean(player.chose_activism, na.rm = TRUE),
    "Avg Payoff"      = mean(round_payoff, na.rm = TRUE),
    "Avg Wage" = mean(wage, na.rm = TRUE),
    "Avg Norm Wage" = mean(norm_wage, na.rm = TRUE),
    "Avg Welfare" = mean(welfare, na.rm = TRUE),
    "Avg Efficiency" = mean(efficiency, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  kbl(digits = 3, caption = "Outcome means by condition - all data") %>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)



df_long_clean %>%
  filter(!is.na(relative_contribution)) %>%
  group_by(treatment_assigned,is_treatment_round) %>%
  rename(Condition = treatment_assigned,
         "Treatment Round" = is_treatment_round) %>%
  summarise(
    n                = n(),
    "Avg Contribution %" = mean(relative_contribution, na.rm = TRUE),
    "Avg Contribution % (SD)"   = sd(relative_contribution, na.rm = TRUE),
    "Avg # of Activists"    = mean(num_activists, na.rm = TRUE),
    "Avg Activism Rate"    = mean(player.chose_activism, na.rm = TRUE),
    "Avg Payoff"      = mean(round_payoff, na.rm = TRUE),
    "Avg Wage" = mean(wage, na.rm = TRUE),
    "Avg Norm Wage" = mean(norm_wage, na.rm = TRUE),
    "Avg Welfare" = mean(welfare, na.rm = TRUE),
    "Avg Efficiency" = mean(efficiency, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  kbl(digits = 3, caption = "Outcome means by condition - complete data only") %>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)


```

```{r questionnaire summary by condition}
#describeBy(df_questionnaire,group = "participant.treatment")

library(gtsummary)
df_questionnaire %>%
  tbl_summary(include = c(questionnaire.1.player.q1_support:questionnaire.1.player.q8_annoyance),
              by = participant.treatment,
              type = list(all_continuous() ~ "continuous2",
                          c(questionnaire.1.player.q3_personal_cost:questionnaire.1.player.q8_annoyance
                            ) ~ "continuous2"),
              statistic = list(all_continuous() ~ c("{mean} ({sd})","{median} [{min},{max}]"),
                               all_categorical() ~ "{n} ({p}%)"),
              #sort = list(everything() ~ "frequency"),
              digits = list(all_continuous() ~ 2,
                            all_categorical() ~ 2)) %>%
  add_n() %>% # add column with total number of non-missing observations
  add_overall() %>%
  add_p(pvalue_fun = ~ style_pvalue(.x, digits = 2))

```

# Main Analysis & Plots

```{r color palette, include=FALSE, warning=FALSE}

# Derive named colors from Tableau_10: roadblock=blue[1], petition=orange[2], demonstration=red[3]
.tab10 <- paletteer::paletteer_d("ggthemes::Tableau_10")

condition_colors <- c(
  "control"       = "gray40",
  "petition"      = as.character(.tab10[2]),
  "demonstration" = as.character(.tab10[3]),
  "roadblock"     = as.character(.tab10[1])
)

treatment_colors <- c(
  "petition"      = as.character(.tab10[2]),
  "demonstration" = as.character(.tab10[3]),
  "roadblock"     = as.character(.tab10[1])
)

table(df_long$treatment_assigned)
df_long$treatment_assigned = relevel(as.factor(df_long$treatment_assigned),"demonstration")

```


Prereg: <https://osf.io/uz3h9/overview>

## H1: Contribution Rate

**Hypothesis:** Contribution(Roadblock) \< Control \< Petition \< Demonstration 
**Actual:** Control* \< Demonstration \< Roadblock \< Petition

### OLS (Main Specification)


```{r h1 ols & emmeans}

#1. OLS: main preregistered model - all groups
model_h1_ols <- lm(relative_contribution ~ condition + period,
                   data = df_long)

# Cluster-robust SEs at group level
vcov_h1_ols <- vcovCL(model_h1_ols, cluster = ~ group_id)
coeftest(model_h1_ols, vcov = vcov_h1_ols)


plot_model(model_h1_ols, 
           show.values=T,show.p=T,  sort.est = T,#colors = condition_colors,
           bpe.style = "dot", value.offset = .2,dot.size = 3,#mdrt.values ="meansd",
                     jitter = 1, value.size = 3,line.size = .5) + theme_classic()



#cleaner version but slightly more conservative results
model_h1_ols_rob <- estimatr::lm_robust(relative_contribution ~ condition + period,
                   data = df_long,
                   clusters = group_id,
                   se_type = "CR2") #because small sample
tab_model(model_h1_ols_rob)

 


#2. estimated marginal means
emm_df <- emmeans(model_h1_ols, pairwise ~ condition, adjust = "holm")
nice_table((emmeans(model_h1_ols, pairwise ~ condition, adjust = "holm")$emmeans))
nice_table((emmeans(model_h1_ols, pairwise ~ condition, adjust = "holm")$contrasts))

# Save pairwise contrasts (Holm-adjusted) for the appendix
h1_contrasts_df <- as.data.frame(
  emmeans(model_h1_ols, pairwise ~ condition, adjust = "holm")$contrasts
) %>%
  transmute(
    Contrast = paste(str_to_title(gsub(" - .*", "", contrast)), "vs.",
                      str_to_title(gsub(".*- ", "", contrast))),
    Estimate = round(estimate, 2),
    SE = round(SE, 2),
    df = round(df, 0),
    t = round(t.ratio, 2),
    p = ifelse(p.value < 0.001, "<0.001", sprintf("%.3f", p.value))
  )
modelsummary::datasummary_df(h1_contrasts_df, output = paper_path("tables", "results", "h1_contrasts.tex"))
strip_table_wrapper(paper_path("tables", "results", "h1_contrasts.tex"))




#3. anova version
model_h1_aov <- aov(relative_contribution ~ condition + as.factor(period),
                   data = df_long)
summary(model_h1_aov)
TukeyHSD(model_h1_aov) #less conservative than emmeans; rb & demo sig relative to control


```

### Tobit (Robustness with Bounded Outcome [0, 1])

```{r h1 tobit}

#results different when tobit model used (demonstration effect becomes neg)

model_h1_tobit <- tobit(relative_contribution ~ condition + period,
                        left  = 0, right = 100,
                        data  = df_long)

vcov_h1_tobit <- vcovCL(model_h1_tobit, cluster = ~ group_id)
coeftest(model_h1_tobit, vcov = vcov_h1_tobit)

nice_table((emmeans(model_h1_tobit, pairwise ~ condition, adjust = "holm")$emmeans))
nice_table((emmeans(model_h1_tobit, pairwise ~ condition, adjust = "holm")$contrasts))


```

```{r h1 (exploratory) with group-based control}

#v2 with treatment_assigned instead of condition 
model_h1b_ols <- lm(relative_contribution ~ treatment_assigned * is_treatment_round + period,
                   data = df_long)
summary(model_h1b_ols)
vcov_h1b_ols <- vcovCL(model_h1b_ols, cluster = ~ group_id)
coeftest(model_h1b_ols, vcov = vcov_h1b_ols)

nice_table((emmeans(model_h1b_ols, pairwise ~ treatment_assigned * is_treatment_round, adjust = "holm")$emmeans))

# Save estimated marginal means (Condition x Phase) for the appendix
emm_h1b_df <- as.data.frame(
  emmeans(model_h1b_ols, pairwise ~ treatment_assigned * is_treatment_round, adjust = "holm")$emmeans
) %>%
  transmute(
    Condition = str_to_title(as.character(treatment_assigned)),
    Phase = if_else(is_treatment_round == 1, "Treatment", "Control"),
    Mean = round(emmean, 2),
    SE = round(SE, 2),
    df = round(df, 1),
    `Lower CI` = round(lower.CL, 2),
    `Upper CI` = round(upper.CL, 2)
  )
modelsummary::datasummary_df(emm_h1b_df, output = paper_path("tables", "results", "emm_h1b.tex"))
strip_table_wrapper(paper_path("tables", "results", "emm_h1b.tex"))

interact_plot(model_h1b_ols, pred = "is_treatment_round",modx = "treatment_assigned",
              #plot.points = T,
              #rug = T,jitter = .1,
              #interval = T,
              robust = T,line.thickness = 2,int.type = "confidence",
              colors = c("#F28E2BFF", "#E15759FF","#4E79A7FF")) + 
  theme_classic()

```

### Figs: Contribution Rate

#### 1a. Main Effects (Preregistered)
```{r emm summary, fig.width=9, fig.height=5}

#650x400

# dynamic Holm-adjusted stars from emm_df (already uses adjust = "holm")
emm_h1_vals <- tidy(emm_df$emmeans) %>% select(condition, emm_est = estimate)

sig_h1 <- tidy(emm_df$contrasts) %>%
  mutate(
    grp1  = str_trim(str_split_fixed(contrast, " - ", 2)[, 1]),
    grp2  = str_trim(str_split_fixed(contrast, " - ", 2)[, 2]),
    stars = case_when(
      adj.p.value < 0.001 ~ "***",
      adj.p.value < 0.01  ~ "**",
      adj.p.value < 0.05  ~ "*",
      TRUE                ~ "ns"
    )
  ) %>%
  filter(stars != "ns") %>%
  left_join(emm_h1_vals, by = c("grp1" = "condition")) %>% rename(e1 = emm_est) %>%
  left_join(emm_h1_vals, by = c("grp2" = "condition")) %>% rename(e2 = emm_est) %>%
  mutate(y_pos = pmax(e1, e2) + 2 + (row_number() - 1) * 3)

ggplot(tidy(emm_df$emmeans), aes(x = reorder(condition, estimate), y = estimate, colour = condition)) +
  geom_point(size = 4) +
  geom_errorbar(aes(ymin = estimate - std.error, ymax = estimate + std.error), width = 0.1) +
  #geom_signif(
  #  comparisons = map2(sig_h1$grp1, sig_h1$grp2, ~ c(.x, .y)),
  #  annotations = sig_h1$stars,
  #  y_position  = sig_h1$y_pos + 5,
  #  tip_length  = 0.01,
  #  vjust       = -0.2,
  #  colour      = "darkgray") +
  scale_colour_manual(values = condition_colors) +
  labs(#title = "Estimated Marginal Means: Contribution Rate by Condition",
       #subtitle = "OLS with cluster-robust SEs (clustered by group)",
       x = "Condition", y = "Contribution Rate (0-100%)",
       colour = NULL) +
  theme_classic(base_size = 12) +
  theme(legend.position = "none") +
  geom_hline(yintercept = 50, colour = "darkgray", alpha = .4, size = 0.8, linetype = "dashed") +
  coord_flip() +
  ylim(c(10, 70))


```

```{r plot contributions, fig.width=8, fig.height=5}

summary(df_long$relative_contribution)

fig_1_1_contribution <- df_long %>%
  ggplot(aes(x = condition, y = relative_contribution, fill = condition)) +
  ggdist::stat_halfeye(alpha = 0.7, adjust = 0.8) +
  geom_boxplot(width = 0.1, outlier.shape = NA, alpha = 0.4) +
  scale_fill_manual(values = condition_colors) +
  labs(#title = "Contribution Rate by Condition",
       x = "Condition", y = "Contribution Rate (0-100%)", fill = NULL) +
  theme_classic(base_size = 12) +
  theme(legend.position = "none") +
  scale_y_continuous(breaks = seq(0, 100, 25), limits = c(0, 120 + nrow(sig_h1) * 8)) +
  coord_cartesian(ylim = c(0, 120 + nrow(sig_h1) * 8))
  #geom_signif(
  #  comparisons = map2(sig_h1$grp1, sig_h1$grp2, ~ c(.x, .y)),
  #  annotations = sig_h1$stars,
  #  y_position  = 105 + (seq_len(nrow(sig_h1)) - 1) * 8,
  #  tip_length  = 0.01,
  #  vjust       = -0.2,
  #  colour      = "darkgray")

fig_1_1_contribution

ggsave(paper_path("figures", "results", "1.1.contribution.png"), fig_1_1_contribution,
       width = 8, height = 5, dpi = 300, bg = "white")

```

#### 1b. Main Effects (Pre-Post)
```{r contribution over time - ctl vs treatment round}

emm_phase <- emmeans(model_h1b_ols, ~ treatment_assigned * is_treatment_round,
                     at = list(is_treatment_round = c(0, 1))) %>%
  tidy() %>%
  mutate(phase = factor(
    if_else(is_treatment_round == 0, "Control Rounds\n(Periods 1–7)", "Treatment Rounds\n(Periods 8–14)"),
    levels = c("Control Rounds\n(Periods 1–7)", "Treatment Rounds\n(Periods 8–14)")
  ))

# pairwise contrasts between conditions at treatment phase only
tx_contrasts <- emmeans(model_h1b_ols, pairwise ~ treatment_assigned,
                        at = list(is_treatment_round = 1),
                        adjust = "holm")$contrasts %>%
  tidy() %>%
  mutate(
    grp1  = str_trim(str_split_fixed(contrast, " - ", 2)[, 1]),
    grp2  = str_trim(str_split_fixed(contrast, " - ", 2)[, 2]),
    stars = case_when(
      adj.p.value < 0.001 ~ "***",
      adj.p.value < 0.01  ~ "**",
      adj.p.value < 0.05  ~ "*",
      TRUE                ~ "ns"
    )
  ) %>%
  filter(stars != "ns")

# y-positions at treatment phase for bracket placement
tx_ypos <- emm_phase %>%
  filter(is_treatment_round == 1) %>%
  select(treatment_assigned, estimate) %>%
  deframe()

# bracket coordinates: vertical segment to the right of x=2, one per sig pair
bracket_data <- tx_contrasts %>%
  mutate(
    y1    = map_dbl(grp1, ~ tx_ypos[.x]),
    y2    = map_dbl(grp2, ~ tx_ypos[.x]),
    x_off = 2 + row_number() * 0.08
  )

p_phase <- ggplot(emm_phase, aes(x = phase, y = estimate, colour = treatment_assigned, group = treatment_assigned)) +
  geom_point(size = 4) +
  geom_line(linewidth = 0.8) +
  geom_errorbar(aes(ymin = estimate - std.error, ymax = estimate + std.error), width = 0.05) +
  scale_colour_manual(values = treatment_colors) +
  labs(#title = "Contribution Rate: Control vs Treatment Rounds",
       x = NULL, y = "Contribution Rate (0-100%)", colour = NULL) +
  theme_classic(base_size = 12)

# significance-star brackets disabled (headers/stars now live in the caption, not the figure)
if (FALSE && nrow(bracket_data) > 0) {
  p_phase <- p_phase +
    geom_segment(data = bracket_data,
                 aes(x = x_off, xend = x_off, y = y1, yend = y2),
                 inherit.aes = FALSE, colour = "grey40", linewidth = 0.4) +
    geom_segment(data = bracket_data,
                 aes(x = x_off - 0.02, xend = x_off, y = y1, yend = y1),
                 inherit.aes = FALSE, colour = "grey40", linewidth = 0.4) +
    geom_segment(data = bracket_data,
                 aes(x = x_off - 0.02, xend = x_off, y = y2, yend = y2),
                 inherit.aes = FALSE, colour = "grey40", linewidth = 0.4) +
    geom_text(data = bracket_data,
              aes(x = x_off + 0.04, y = (y1 + y2) / 2, label = stars),
              inherit.aes = FALSE, colour = "grey40", size = 4)
}

fig1_panelA <- p_phase +
  coord_cartesian(xlim = c(1, 2), ylim = c(0, 60), clip = "off") +
  theme(legend.position = "top")

fig1_panelA

ggsave(paper_path("figures", "results", "1.2.contributionprepost.png"), fig1_panelA,
       width = 8, height = 5, dpi = 300, bg = "white")


```

#### 1c. Main Effects (Over Time)
```{r contribution over time, fig.width=9, fig.height=5}

fig1_panelB <- df_long %>%
  #filter(!is.na(relative_contribution)) %>%
  group_by(period, treatment_assigned) %>%
  summarise(mean_rc = mean(relative_contribution, na.rm = TRUE),
            se_rc   = sd(relative_contribution, na.rm = TRUE) / sqrt(n()),
            .groups = "drop") %>%
  ggplot(aes(x = period, y = mean_rc, colour = treatment_assigned, group = treatment_assigned)) +
  geom_vline(xintercept = 7.5, linetype = "dashed", colour = "grey50") +
  geom_line(linewidth = 1) +
  geom_ribbon(aes(ymin = mean_rc - se_rc, ymax = mean_rc + se_rc, fill = treatment_assigned),
              alpha = 0.15, colour = NA) +
  scale_colour_manual(values = treatment_colors) +
  scale_fill_manual(values = treatment_colors) +
  annotate("text", x = 3.5, y = Inf, label = "Control Period", vjust = 1.5,
           colour = "grey40", size = 3) +
  annotate("text", x = 11,  y = Inf, label = "Treatment Period",   vjust = 1.5,
           colour = "grey40", size = 3) +
  labs(#title = "Mean Relative Contribution Over Time",
       x = "Period", y = "Mean Relative Contribution",
       colour = NULL, fill = NULL) +
  theme_classic(base_size = 12) +
  ylim(c(0,60)) +
  #scale_x_continuous(breaks = c(2,4,8,6,10,12,14)) +
  theme(legend.position = "top")

fig1_panelB

ggsave(paper_path("figures", "results", "1.4.contributionovertime.png"), fig1_panelB,
       width = 9, height = 5, dpi = 300, bg = "white")


df_long %>%
  filter(!is.na(relative_contribution)) %>%
  group_by(period, condition) %>%
  summarise(mean_rc = mean(relative_contribution, na.rm = TRUE),
            se_rc   = sd(relative_contribution, na.rm = TRUE) / sqrt(n()),
            .groups = "drop") %>%
  ggplot(aes(x = period, y = mean_rc, colour = condition, group = condition)) +
  geom_vline(xintercept = 7.5, linetype = "dashed", colour = "grey50") +
  geom_line(linewidth = 1) +
  geom_ribbon(aes(ymin = mean_rc - se_rc, ymax = mean_rc + se_rc, fill = condition),
              alpha = 0.15, colour = NA) +
  scale_colour_manual(values = condition_colors) +
  scale_fill_manual(values = condition_colors) +
  annotate("text", x = 3.5, y = Inf, label = "Control Period", vjust = 1.5,
           colour = "grey40", size = 3) +
  annotate("text", x = 11,  y = Inf, label = "Treatment Period",   vjust = 1.5,
           colour = "grey40", size = 3) +
  labs(#title = "Mean Relative Contribution Over Time",
       x = "Period", y = "Mean Relative Contribution",
       colour = NULL, fill = NULL) +
  theme_classic(base_size = 12) +
  #scale_x_continuous(breaks = c(2,4,8,6,10,12,14)) +
  theme(legend.position = "top")

```

#### 1d. Figure 1: Combined Panels (Control vs. Treatment Round + Contribution over Period)
```{r figure 1 combined, fig.width=8, fig.height=9}

# panel B's legend keys (line + ribbon fill) don't match panel A's (point + line),
# so guides = "collect" places them side by side instead of merging; suppress panel
# B's legend entirely and let panel A's be the one shared legend for the figure
figure1 <- (fig1_panelA / (fig1_panelB + guides(colour = "none", fill = "none"))) +
  plot_annotation(tag_levels = "A") +
  plot_layout(guides = "collect") &
  theme(legend.position = "top")

figure1

ggsave(paper_path("figures", "results", "figure1.png"), figure1,
       width = 8, height = 9, dpi = 300, bg = "white")

```

## H2: Activism Turnout

**Hypothesis:** Turnout(Roadblock) \< Demonstration \< Petition

Only available for treatment rounds (>=7) from period t=8 onwards (lagged predictors require t-1 data).

```{r h2 data}
df_activism <- df_long %>%
  filter(
    period >= 8,
    !is.na(chose_activism),
    !is.na(others_contribution_lag),
    !is.na(NegDev_lag),
    !is.na(PosDev_lag)
  )

cat("Observations for activism models:", nrow(df_activism))
```

TO DO: distribution kinda weird & varied by condition so figure out which model best accounts for this
```{r summary}

summary(df_activism$activism_rate)
table(df_activism$player.chose_activism,df_activism$condition)

summary(df_activism$num_activists)
table(df_activism$num_activists,df_activism$condition) #in most cases people didn't demonstrate, a good bit often chose roadblocking

```

### NUMERIC VERSION
```{r ols with full df}
#if we use num_activists/activist rate it duplicates for each group member so just use the group level version

#pure - all groups
model_h2_ols <- lm(player.chose_activism * 100 ~ condition + others_contribution_lag + NegDev_lag + PosDev_lag + period,
                   data = df_long)

nice_table((emmeans(model_h2_ols, pairwise ~ condition, adjust = "holm")$emmeans))
nice_table((emmeans(model_h2_ols, pairwise ~ condition, adjust = "holm")$contrasts))


# Cluster-robust SEs 
coeftest(model_h2_ols, vcov = vcovCL(model_h2_ols, cluster = ~ group_id))


# Reference = demonstration, not control (the project default): control has zero
# valid rows for this outcome (no control-period activism turnout to speak of), so a
# control reference leaves the model rank-deficient and R silently drops a different
# dummy instead.
model_h2_ols_rob <- estimatr::lm_robust(player.chose_activism ~ condition + others_contribution_lag + NegDev_lag + PosDev_lag + period,
                   data = df_long %>% mutate(condition = relevel(factor(condition), ref = "demonstration")),
                   clusters = group_id,
                   se_type = "CR2") #because small sample
tab_model(model_h2_ols_rob)

# Save lagged-controls model for the paper (Table \ref{tab:model_lag})
save_reg_table(list("Activism Turnout" = model_h2_ols_rob),
               path = paper_path("tables", "results", "model_lag.tex"))

# Main-table version: condition + period only, no lagged controls -- matches
# the H1/H3 specification and feeds the "Activism Turnout" column of
# tables/results/mainreg.tex. The lagged specification above is the dedicated
# test of the Fehr (2000) reciprocity hypothesis (Table \ref{tab:model_lag}).
# Reference = demonstration, not control, for the same reason as model_h2_ols_rob above.
model_h2_main_rob <- estimatr::lm_robust(player.chose_activism ~ condition + period,
                   data = df_long %>% mutate(condition = relevel(factor(condition), ref = "demonstration")),
                   clusters = group_id,
                   se_type = "CR2")
tab_model(model_h2_main_rob)


```


#### Group-Level Activism Turnout
```{r h2 group level regressions}

df_activism_group <- df_long %>%
  filter(is_treatment_round == 1) %>%
  group_by(group_id, period, condition) %>%
  summarise(
    turnout_rate = mean(player.chose_activism, na.rm = TRUE) * 100,
    n_activists  = sum(num_activists, na.rm = TRUE),
    
    NegDev_lag = mean(NegDev_lag, na.rm = TRUE),
    PosDev_lag = mean(PosDev_lag, na.rm = TRUE), #pos & neg the same when aggregated? it drops posdev
    others_contribution_lag = mean(others_contribution_lag, na.rm = TRUE),
    
    group_size   = n(),
    .groups      = "drop")

model_h2_group <- estimatr::lm_robust(turnout_rate ~ condition + others_contribution_lag + NegDev_lag + PosDev_lag + period,
                   data = df_activism_group,
                   clusters = group_id,
                   se_type = "CR2") #because small sample

tab_model(model_h2_ols_rob)
```


### BINARY VERSION (Logistic/Main Specification)
```{r h2 logit (prereg)}

model_h2_logit <- glm(player.chose_activism ~ condition + others_contribution_lag + NegDev_lag + PosDev_lag + period,
                      family = binomial(link = "logit"), data = df_activism)
tab_model(model_h2_logit)

vcov_h2_logit <- vcovCL(model_h2_logit, cluster = ~ group_id)
coeftest(model_h2_logit, vcov = vcov_h2_logit)


emm_h2_logit <- emmeans(model_h2_logit, "condition", vcov. = vcov_h2_logit)
emm_h2_logit

```

### Probit (Robustness)

```{r h2 probit}
model_h2_probit <- glm(player.chose_activism ~ condition + others_contribution_lag +
    NegDev_lag + PosDev_lag + period,
  family = binomial(link = "probit"), data = df_activism)

vcov_h2_probit <- vcovCL(model_h2_probit, cluster = ~ group_id)
coeftest(model_h2_probit, vcov = vcov_h2_probit)

```


### Figs: Turnout

#### 2a. Main Effects (Preregistered)
```{r turnout by condition, fig.width=8, fig.height=5}

fig_2_1_activismrate <- df_activism_group %>%
  ggplot(aes(x = condition, y = turnout_rate, fill = condition)) +
  ggdist::stat_halfeye(alpha = 0.7, adjust = 0.8) +
  geom_boxplot(width = 0.1, outlier.shape = NA, alpha = 0.4) +
  scale_fill_manual(values = condition_colors) +
  labs(#title = "Activism Turnout Rate by Condition",
       #subtitle = "Treatment rounds only (periods 8–14)",
       x = "Condition", y = "Turnout Rate (%)", fill = NULL) +
  theme_classic(base_size = 12) +
  theme(legend.position = "none") +
  geom_hline(yintercept = 50, colour = "darkgray", alpha = .4, size = 0.8, linetype = "dashed")

fig_2_1_activismrate

ggsave(paper_path("figures", "results", "2.1.activismrate.png"), fig_2_1_activismrate,
       width = 8, height = 5, dpi = 300, bg = "white")

```

#### 2b. Main Effects (Over Time)
```{r turnout over time, fig.width=9, fig.height=5}

#on group level data
fig_2_1_activismrateovertime <- df_activism_group %>%
  group_by(period, condition) %>%
  summarise(
    mean_turnout = mean(turnout_rate, na.rm = TRUE),
    se_turnout   = sd(turnout_rate, na.rm = TRUE) / sqrt(n()),
    .groups = "drop") %>%
  ggplot(aes(x = period, y = mean_turnout, colour = condition, group = condition)) +
  geom_line(linewidth = 1) +
  geom_ribbon(aes(ymin = mean_turnout - se_turnout, ymax = mean_turnout + se_turnout,
                  fill = condition), alpha = 0.15, colour = NA) +
  scale_colour_manual(values = treatment_colors) +
  scale_fill_manual(values = treatment_colors) +
  labs(#title = "Activism Turnout Rate Over Time",
       #subtitle = "Treatment rounds only (periods 8-14)",
       x = "Period", y = "Mean Turnout Rate (%)", colour = NULL, fill = NULL) +
  theme_classic(base_size = 12) +
  xlim(c(8, 14))

fig_2_1_activismrateovertime

ggsave(paper_path("figures", "results", "2.1.activismrateovertime.png"), fig_2_1_activismrateovertime,
       width = 9, height = 5, dpi = 300, bg = "white")

#on individual-level data
df_long %>%
  group_by(period, condition) %>%
  summarise(
    mean_turnout = mean(player.chose_activism, na.rm = TRUE) *100,
    se_turnout   = sd(player.chose_activism, na.rm = TRUE) / sqrt(n()) * 100,
    .groups = "drop") %>%
  ggplot(aes(x = period, y = mean_turnout, colour = condition, group = condition)) +
  geom_line(linewidth = 1) +
  geom_ribbon(aes(ymin = mean_turnout - se_turnout, ymax = mean_turnout + se_turnout,
                  fill = condition), alpha = 0.15, colour = NA) +
  scale_colour_manual(values = treatment_colors) +
  scale_fill_manual(values = treatment_colors) +
  labs(#title = "Activism Turnout Rate Over Time",
       #subtitle = "Treatment rounds only (periods 8-14)",
       x = "Period", y = "Mean Turnout Rate (%)", colour = NULL, fill = NULL) +
  theme_classic(base_size = 12) +
  xlim(c(8, 14)) 
```


```{r contribution & turnout over time}

ggerrorplot(df_long, 
            x = "treatment_assigned", y = c("relative_contribution","activism_rate"),
            combine = T,
            size = 1,
            add.params = list(),legend = "top",
            position = position_jitter(.2), alpha = .8, orientation =  "horiz", add = "mean_se",
            na.rm = T, sort.val = "asc") +
  geom_hline(yintercept = 0, colour = "darkgray", alpha = .4, size = 0.8, linetype = "dashed") + 
  labs(x = "Condition", y = "Rate"#, title = "Contribution & Activism Rate by Condition"
       ) +  # Custom legend title
  theme_classic()  +
  theme(legend.title = element_text(size = 10, face = "bold"))


```

## H3: Welfare

**Hypothesis:** Welfare(Roadblock) \< Demonstration \< Petition

### Welfare = total group earnings per period.

```{r h3 welfare ols}
summary(df_long$welfare)

model_h3 <- lm(welfare ~ condition + period, data = df_long)
vcov_h3   <- vcovCL(model_h3, cluster = ~ group_id)
coeftest(model_h3, vcov = vcov_h3)

nice_table((emmeans(model_h3, pairwise ~ condition, adjust = "holm")$emmeans))
nice_table((emmeans(model_h3, pairwise ~ condition, adjust = "holm")$contrasts))

model_h3_ols_rob <- estimatr::lm_robust(welfare ~ condition + period,
                   data = df_long,
                   clusters = group_id,
                   se_type = "CR2") #because small sample
tab_model(model_h3_ols_rob)



model_h3norm <- lm(norm_welfare ~ condition + period, data = df_long)
vcov_h3norm   <- vcovCL(model_h3norm, cluster = ~ group_id)
coeftest(model_h3norm, vcov = vcov_h3norm)

nice_table((emmeans(model_h3norm, pairwise ~ condition, adjust = "holm")$emmeans))
nice_table((emmeans(model_h3norm, pairwise ~ condition, adjust = "holm")$contrasts))


#nothing sig
model_h3norm_ols_rob <- estimatr::lm_robust(norm_welfare ~ condition + period,
                   data = df_long,
                   clusters = group_id,
                   se_type = "CR2") #because small sample
tab_model(model_h3norm_ols_rob)

```


### Figs: Welfare 
#### 3a. Main Effects (Preregistered)
```{r h3 plot pooled condition, fig.width=7, fig.height=4}

# 1a. by condition
df_long %>%
  filter(!is.na(welfare) & welfare > 0) %>%
  ggplot(aes(x = condition, y = welfare, fill = condition)) +
  geom_boxplot(alpha = 0.7) +
  geom_jitter(width = 0.1, alpha = 0.1, size = 1.5) +
  scale_fill_manual(values = condition_colors) +
  labs(#title = "Group Welfare by Condition",
       x = "Condition", y = "Total Group Earnings (per period)",
       fill = NULL) +
  theme_classic(base_size = 12) +
  theme(legend.position = "none")


# 1b. normalized welfare
df_long %>%
  filter(!is.na(norm_welfare) & norm_welfare > 0) %>%
  ggplot(aes(x = condition, y = norm_welfare, fill = condition)) +
  geom_boxplot(alpha = 0.7) +
  geom_jitter(width = 0.1, alpha = 0.1, size = 1.5) +
  scale_fill_manual(values = condition_colors) +
  labs(#title = "Group Welfare by Condition (Time-Normalized)",
       x = "Condition", y = "Welfare per Second (per period)",
       fill = NULL) +
  theme_classic(base_size = 12) +
  theme(legend.position = "none")

```

```{r h3 before vs after, fig.width=7, fig.height=4}

# OLS models mirroring model_h1b_ols structure for phase comparisons
model_h3b_ols     <- lm(welfare      ~ treatment_assigned * is_treatment_round + period,
                         data = df_long %>% filter(!is.na(welfare) & welfare > 0))
model_h3bnorm_ols <- lm(norm_welfare ~ treatment_assigned * is_treatment_round + period,
                         data = df_long %>% filter(!is.na(norm_welfare) & norm_welfare > 0))

# helper: EMMs at both phases + between-group brackets at treatment phase
phase_plot <- function(model, y_label, title_label) {
  emm_ph <- emmeans(model, ~ treatment_assigned * is_treatment_round,
                    at = list(is_treatment_round = c(0, 1))) %>%
    tidy() %>%
    mutate(phase = factor(
      if_else(is_treatment_round == 0, "Control Rounds\n(Periods 1–7)", "Treatment Rounds\n(Periods 8–14)"),
      levels = c("Control Rounds\n(Periods 1–7)", "Treatment Rounds\n(Periods 8–14)")
    ))

  tx_ypos <- emm_ph %>%
    filter(is_treatment_round == 1) %>%
    select(treatment_assigned, estimate) %>%
    deframe()

  brk <- emmeans(model, pairwise ~ treatment_assigned,
                 at = list(is_treatment_round = 1),
                 adjust = "holm")$contrasts %>%
    tidy() %>%
    mutate(
      grp1  = str_trim(str_split_fixed(contrast, " - ", 2)[, 1]),
      grp2  = str_trim(str_split_fixed(contrast, " - ", 2)[, 2]),
      stars = case_when(
        adj.p.value < 0.001 ~ "***",
        adj.p.value < 0.01  ~ "**",
        adj.p.value < 0.05  ~ "*",
        TRUE                ~ "ns"
      )
    ) %>%
    filter(stars != "ns") %>%
    mutate(
      y1    = map_dbl(grp1, ~ tx_ypos[.x]),
      y2    = map_dbl(grp2, ~ tx_ypos[.x]),
      x_off = 2 + row_number() * 0.08
    )

  p <- ggplot(emm_ph, aes(x = phase, y = estimate, colour = treatment_assigned, group = treatment_assigned)) +
    geom_point(size = 3) +
    geom_line(linewidth = 0.8) +
    geom_errorbar(aes(ymin = estimate - std.error, ymax = estimate + std.error), width = 0.1) +
    scale_colour_manual(values = treatment_colors) +
    labs(#title = title_label,
         x = NULL, y = y_label, colour = NULL) +
    theme_classic(base_size = 12) +
    theme(legend.position = "top")

  # significance-star brackets disabled (headers/stars now live in the caption, not the figure)
  if (FALSE && nrow(brk) > 0) {
    p <- p +
      geom_segment(data = brk, aes(x = x_off, xend = x_off, y = y1, yend = y2),
                   inherit.aes = FALSE, colour = "grey40", linewidth = 0.4) +
      geom_segment(data = brk, aes(x = x_off - 0.02, xend = x_off, y = y1, yend = y1),
                   inherit.aes = FALSE, colour = "grey40", linewidth = 0.4) +
      geom_segment(data = brk, aes(x = x_off - 0.02, xend = x_off, y = y2, yend = y2),
                   inherit.aes = FALSE, colour = "grey40", linewidth = 0.4) +
      geom_text(data = brk, aes(x = x_off + 0.04, y = (y1 + y2) / 2, label = stars),
                inherit.aes = FALSE, colour = "grey40", size = 4) +
      coord_cartesian(clip = "off")
  }
  p
}

# raw welfare
fig_3_3_welfareprepost <- phase_plot(model_h3b_ols, "EMM Group Welfare",
           "Group Welfare: Control vs Treatment Rounds by Assigned Condition")
fig_3_3_welfareprepost

ggsave(paper_path("figures", "results", "3.3.welfareprepost.png"), fig_3_3_welfareprepost,
       width = 7, height = 4, dpi = 300, bg = "white")

# time-normalized welfare
phase_plot(model_h3bnorm_ols, "EMM Welfare per Second",
           "Welfare (Time-Normalized): Control vs Treatment Rounds by Assigned Condition")

```

#### 3b. Main Effects (Over Time)
```{r plot over time, fig.width=9, fig.height=5}

# 3. Welfare over time by condition
df_long %>%
  filter(!is.na(welfare) & welfare > 0) %>%
  group_by(period, condition) %>%
  summarise(
    mean_welfare = mean(welfare, na.rm = TRUE),
    se_welfare   = sd(welfare, na.rm = TRUE) / sqrt(n()),
    .groups = "drop"
  ) %>%
  ggplot(aes(x = period, y = mean_welfare, colour = condition, group = condition)) +
  geom_vline(xintercept = 7.5, linetype = "dashed", colour = "grey50") +
  geom_line(linewidth = 0.8) +
  geom_ribbon(aes(ymin = mean_welfare - se_welfare, ymax = mean_welfare + se_welfare, fill = condition),
              alpha = 0.15, colour = NA) +
  annotate("text", x = 3.5, y = Inf, label = "Control Period", vjust = 1.5,
           colour = "grey40", size = 3) +
  annotate("text", x = 11,  y = Inf, label = "Treatment Period",   vjust = 1.5,
           colour = "grey40", size = 3) +
  scale_colour_manual(values = condition_colors) +
  scale_fill_manual(values = condition_colors) +
  labs(#title = "Group Welfare Over Time by Condition",
       x = "Period", y = "Mean Group Welfare", colour = NULL, fill = NULL) +
  theme_classic(base_size = 12) +
  xlim(c(1, 14))  +
    theme(legend.position = "top")


# 4. Welfare over time by treatment_assigned
df_long %>%
  filter(!is.na(welfare) & welfare > 0) %>%
  group_by(period, treatment_assigned) %>%
  summarise(
    mean_welfare = mean(welfare, na.rm = TRUE),
    se_welfare   = sd(welfare, na.rm = TRUE) / sqrt(n()),
    .groups = "drop"
  ) %>%
  ggplot(aes(x = period, y = mean_welfare, colour = treatment_assigned, group = treatment_assigned)) +
  geom_vline(xintercept = 7.5, linetype = "dashed", colour = "grey50") +
  geom_line(linewidth = 0.8) +
  geom_ribbon(aes(ymin = mean_welfare - se_welfare, ymax = mean_welfare + se_welfare, fill = treatment_assigned),
              alpha = 0.15, colour = NA) +
  annotate("text", x = 3.5, y = Inf, label = "Control Period", vjust = 1.5,
           colour = "grey40", size = 3) +
  annotate("text", x = 11,  y = Inf, label = "Treatment Period",   vjust = 1.5,
           colour = "grey40", size = 3) +
  scale_colour_manual(values = treatment_colors) +
  scale_fill_manual(values = treatment_colors) +
  labs(#title = "Group Welfare Over Time by Assigned Condition",
       x = "Period", y = "Mean Group Welfare (Points)", colour = NULL, fill = NULL) +
  theme_classic(base_size = 12) +
  xlim(c(1, 14)) +
    theme(legend.position = "top")

```


# Exploratory Analysis

## Efficiency

Efficiency = total group payoff / potential payoff under full cooperation.

```{r efficiency}
summary(df_long$efficiency)
summary(df_long$welfare)
summary(df_long$welfare_fullcoop)

model_efficiency <- lm(efficiency ~ condition + period, data = df_long)
vcov_efficiency  <- vcovCL(model_efficiency, cluster = ~ group_id)
coeftest(model_efficiency, vcov = vcov_efficiency)

# Cluster-robust version (CR2), matching the other mainreg models -- feeds the
# "Efficiency" column of tables/results/mainreg.tex
model_efficiency_rob <- estimatr::lm_robust(efficiency ~ condition + period,
                   data = df_long,
                   clusters = group_id,
                   se_type = "CR2")
tab_model(model_efficiency_rob)

model_efficiency_emm <- emmeans(model_efficiency, pairwise ~ condition, adjust = "holm")

nice_table((emmeans(model_efficiency, pairwise ~ condition, adjust = "holm")$emmeans))
nice_table((emmeans(model_efficiency, pairwise ~ condition, adjust = "holm")$contrasts))


plot_model(model_efficiency)

```

#### 4. Fig: Efficiency 
```{r efficiency fig}

model_efficiency_emm$contrasts

# control-only comparisons (Holm-adjusted)
emm_eff_vals <- tidy(model_efficiency_emm$emmeans) %>% select(condition, emm_est = estimate)

sig_eff <- tidy(model_efficiency_emm$contrasts) %>%
  mutate(
    grp1  = str_trim(str_split_fixed(contrast, " - ", 2)[, 1]),
    grp2  = str_trim(str_split_fixed(contrast, " - ", 2)[, 2]),
    stars = case_when(
      adj.p.value < 0.001 ~ "***",
      adj.p.value < 0.01  ~ "**",
      adj.p.value < 0.05  ~ "*",
      TRUE                ~ "ns"
    )
  ) %>%
  filter(stars != "ns", grp1 == "control" | grp2 == "control") %>%
  left_join(emm_eff_vals, by = c("grp1" = "condition")) %>% rename(e1 = emm_est) %>%
  left_join(emm_eff_vals, by = c("grp2" = "condition")) %>% rename(e2 = emm_est) %>%
  mutate(y_pos = pmax(e1, e2) + 2 + (row_number() - 1) * 3)

fig_5_efficiency <- ggplot(tidy(model_efficiency_emm$emmeans), aes(x = reorder(condition, estimate), y = estimate, colour = condition)) +
  geom_point(size = 4) +
  geom_errorbar(aes(ymin = estimate - std.error, ymax = estimate + std.error), width = 0.1) +
  #geom_signif(
  #  comparisons  = map2(sig_eff$grp1, sig_eff$grp2, ~ c(.x, .y)),
  #  annotations  = sig_eff$stars,
  #  y_position   = sig_eff$y_pos,
  #  tip_length   = 0.01,
  #  vjust        = -0.2,
  #  colour       = "darkgray") +
  scale_colour_manual(values = condition_colors) +
  labs(#title = "Estimated Marginal Means: Efficiency by Condition",
       #subtitle = "OLS with cluster-robust SEs (relative to control)",
       x = "Condition", y = "Efficiency (0-100%)",
       colour = NULL) +
  theme_classic(base_size = 12) +
  theme(legend.position = "none") +
  coord_flip() +
  ylim(c(60, 90))

fig_5_efficiency

ggsave(paper_path("figures", "results", "5.efficiency.png"), fig_5_efficiency,
       width = 7, height = 5, dpi = 300, bg = "white")

```


## Activist vs. Non-Activist Contributions

```{r activist vs nonactivist summary stats}
#interesting cuz activists earned less on avg (not as good at RET typically) & contributed more but got closer to non-activist income levels (esp in petition condition) - are they really prosocial or just disadvantaged - or is 


df_long %>%
  dplyr::filter(is_treatment_round == 1) %>%
  dplyr::mutate(activist = factor(player.chose_activism, labels = c("Non-Activist", "Activist"))) %>%
  dplyr::group_by(condition, player.chose_activism) %>%
  dplyr::summarise(n = n(),
    "Wage" = mean(wage, na.rm = TRUE),
    "Norm Wage" = mean(wage, na.rm = TRUE) / mean(actual_effort_time, na.rm = TRUE),
    
    'Avg Contribution %' = mean(relative_contribution, na.rm = TRUE),
    "Contribution % (SD)" = sd(relative_contribution, na.rm = TRUE),
    "Contribution" = mean(contribution, na.rm = TRUE),
    
    "Avg Payoff" = mean(round_payoff, na.rm = TRUE),
    "Avg Added Wage" = mean(round_payoff, na.rm = TRUE) - mean(wage, na.rm = TRUE),
    .groups = "drop") %>%
  kbl(digits = 2, caption = "Contribution Rate: Activists vs. Non-activists") %>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)



df_long %>%
  dplyr::filter(is_treatment_round == 1) %>%
  dplyr::mutate(activist = factor(player.chose_activism, labels = c("Non-Activist", "Activist"))) %>%
  dplyr::group_by(player.chose_activism) %>%
  dplyr::summarise(n = n(),
    "Wage" = mean(wage, na.rm = TRUE),
    "Norm Wage" = mean(wage, na.rm = TRUE) / mean(actual_effort_time, na.rm = TRUE),
    
    'Avg Contribution %' = mean(relative_contribution, na.rm = TRUE),
    "Contribution % (SD)" = sd(relative_contribution, na.rm = TRUE),
    "Contribution" = mean(contribution, na.rm = TRUE),
    
    "Avg Payoff" = mean(round_payoff, na.rm = TRUE),
    "Avg Added Wage" = mean(round_payoff, na.rm = TRUE) - mean(wage, na.rm = TRUE),
    .groups = "drop") %>%
  kbl(digits = 2, caption = "Contribution Rate: Activists vs. Non-activists") %>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)

```

```{r activist vs nonactivist model}

summary(df_long$player.chose_activism)

model_activist <-estimatr::lm_robust(relative_contribution ~ condition * (player.chose_activism) + period,
                   data = df_long,
                   clusters = group_id,
                   se_type = "CR2")
tab_model(model_activist)


```

##### 5. Fig: Activism Choice & Contribution
```{r chose vs didnt & contribution}
#can't run meaningful model here because a lot of people didn't choose activism in some conditions; in RB always at least 1 activist
hactiv.lmer <- lmer(relative_contribution ~ treatment_assigned * player.chose_activism + period + (1 | group_id) + (1 | participant.code), 
           data = df_long) 
tab_model(hactiv.lmer)

fig_4_activism_cont <- interact_plot(hactiv.lmer, modx = treatment_assigned, pred = player.chose_activism,
              modx.labels = c("Demonstration", "Petition", "Roadblock"),
              legend.main = "Condition",
              plot.points = T, rug = T, jitter = .1, linearity.check = FALSE, point.alpha = .3,
              #interval = T,
              robust = T, line.thickness = 2, int.type = "confidence",
              colors = c("#4E79A7FF", "#E15759FF","#F28E2BFF")) +
  labs(#title = "Interaction Effects: Contribution Rates & Activism Decision",
       x = "If Player Chose Activism", y = "Contribution Rate (0-100%)",
       colour = "Condition") +
  theme_classic(base_size = 12) +
  theme(legend.position = "top") +
  geom_hline(yintercept = 50, colour = "darkgray", alpha = .4, size = 0.8, linetype = "dashed")

fig_4_activism_cont

ggsave(paper_path("figures", "results", "4.activism&cont.png"), fig_4_activism_cont,
       width = 7, height = 5, dpi = 300, bg = "white")

```


## H1-3: Linear Mixed Effects Models for all Outcomes

### Main Effects Model
```{r LME version - main effects (control vs treatments)}

h1.lmer <- lmer(relative_contribution ~ condition + period + (1 | group_id) + (1 | participant.code), 
           data = df_long)
summary(h1.lmer)

h2.lmer <- lmer(player.chose_activism * 100 ~ condition + others_contribution_lag + NegDev_lag + PosDev_lag + period + (1 | group_id) + (1 | participant.code), 
           data = df_long)
summary(h2.lmer)

h3.lmer <- lmer(welfare ~ condition + period + (1 | group_id) + (1 | participant.code), 
           data = df_long)
summary(h3.lmer)

h3b.lmer <- lmer(norm_welfare ~ condition + period + (1 | group_id) + (1 | participant.code), 
           data = df_long)
summary(h3b.lmer)

library(performance)
icc(h1.lmer)

library(corrplot)
correls <- cor(scale(df_long %>% dplyr::select(wage,contribution,relative_contribution,
                                               player.chose_activism,welfare,norm_welfare,added_wage)),
    use = "complete.obs")
correls
tab_corr(correls,show.p = TRUE,p.numeric = T)

corrplot(correls,type = 'lower',diag = F)

```

### Interaction Models (Primary)
```{r main alt model - interaction version to account for baseline differences across conditions}

df_long %>%
  group_by(treatment_assigned,is_treatment_round) %>%
  summarise(mean_contrib = mean(relative_contribution),
            n = n()/4)


#1. CONTRIBUTION RATE
h1.1.lmer <- lmer(relative_contribution ~ treatment_assigned * is_treatment_round + period + 
                    (1 | group_id) + (1 | participant.code), 
           data = df_long)
summary(h1.1.lmer)

nice_table(emmeans(h1.1.lmer, pairwise ~ treatment_assigned,
                         at = list(is_treatment_round = 1), adjust = "holm")$contrasts)


    #alt plots
plot_model(h1.1.lmer,
           type = "int",
           show.values=TRUE,show.p=T,order.terms = T) + coord_flip()


interact_plot(h1.1.lmer,modx = treatment_assigned, pred = is_treatment_round,
              plot.points = T,
              #rug = T,
              jitter = .1,
              point.size = .5,
              #interval = T,
              robust = T,line.thickness = 2,int.type = "confidence",
              colors = treatment_colors) + 
  theme_classic()


ggerrorplot(df_long, 
            x = "treatment_assigned", y = "relative_contribution", 
            #color = "is_treatment_round",
            combine = T,
            size = 1,legend = "top",
            position = position_jitter(.2), alpha = .8, orientation =  "horiz", add = "mean_se",
            na.rm = T, sort.val = "asc") +
  geom_hline(yintercept = 50, colour = "darkgray", alpha = .4, size = 0.8, linetype = "dashed") + 
  labs(x = "AI Feedback Type", y = "Denialism Change (pp)"#, title = "Conversation effects by condition",
       #color = "is_treatment_round", fill = "is_treatment_round"
       ) +  # Custom legend title
  theme_classic()  +
  theme(legend.title = element_text(size = 10, face = "bold"))





#2. ACTIVISM TURNOUT
h2.1.lmer <- lmer(player.chose_activism ~ treatment_assigned + others_contribution_lag + NegDev_lag + PosDev_lag + period +
                    (1 | group_id) + (1 | participant.code),
           data = df_long) #is_treatment_round = 1 only
summary(h2.1.lmer)
nice_table(pairs(emmeans(h2.1.lmer, pairwise ~ treatment_assigned, adjust = "holm")$contrasts))

# Main-table version: treatment_assigned + period only, no lagged controls -- matches
# the period-only H2 specification used in mainreg.tex; feeds the "Activism Turnout"
# column of tables/results/lmer-interaction.tex
h2.1_main.lmer <- lmer(player.chose_activism ~ treatment_assigned + period +
                    (1 | group_id) + (1 | participant.code),
           data = df_long)
summary(h2.1_main.lmer)




#3. WELFARE & EFFICIENCY
h3.1.lmer <- lmer(welfare ~ treatment_assigned * is_treatment_round + period + (1 | group_id) + (1 | participant.code), 
           data = df_long)
summary(h3.1.lmer)
nice_table(pairs(emmeans(h3.1.lmer, pairwise ~ is_treatment_round:treatment_assigned, adjust = "holm")$contrasts))


h3.1b.lmer <- lmer(norm_welfare ~ treatment_assigned * is_treatment_round + period + (1 | group_id) + (1 | participant.code), 
           data = df_long)
summary(h3.1b.lmer)
nice_table(pairs(emmeans(h3.1b.lmer, pairwise ~ is_treatment_round:treatment_assigned, adjust = "holm")$contrasts))


heff.lmer <- lmer(efficiency ~ treatment_assigned * is_treatment_round + period + (1 | group_id) + (1 | participant.code), 
           data = df_long)
summary(heff.lmer)
#nice_table(pairs(emmeans(heff.lmer, pairwise ~ is_treatment_round:treatment_assigned, adjust = "holm")$emmeans))

```

```{r with clean df}

h1.1clean.lmer <- lmer(relative_contribution ~ treatment_assigned * is_treatment_round + period + 
                    (1 | group_id) + (1 | participant.code), 
           data = df_long_clean)

h2.1clean.lmer <- lmer(player.chose_activism ~ treatment_assigned + others_contribution_lag + NegDev_lag + PosDev_lag + period +
                    (1 | group_id) + (1 | participant.code),
           data = df_long_clean) #is_treatment_round = 1 only

# Main-table version: treatment_assigned + period only, no lagged controls -- matches
# h2.1_main.lmer; feeds the "Activism Turnout" column of tables/results/lmer-interaction-comp.tex
h2.1clean_main.lmer <- lmer(player.chose_activism ~ treatment_assigned + period +
                    (1 | group_id) + (1 | participant.code),
           data = df_long_clean)

h3.1clean.lmer <- lmer(welfare ~ treatment_assigned * is_treatment_round + period + (1 | group_id) + (1 | participant.code),
           data = df_long_clean)

h3.1bclean.lmer <- lmer(norm_welfare ~ treatment_assigned * is_treatment_round + period + (1 | group_id) + (1 | participant.code),
           data = df_long_clean)

heffclean.lmer <- lmer(efficiency ~ treatment_assigned * is_treatment_round + period + (1 | group_id) + (1 | participant.code),
           data = df_long_clean)


```

### Tables: All Models
#### Main Preregistered Models (OLS)
```{r clean regression table - preregistered models}

#main preregisted effects, compared to control period
#H2 (Activism Turnout) uses model_h2_main_rob: condition + period only, matching
#the period-only specification of the other three outcomes
tab_model(model_h1_ols_rob,model_h2_main_rob,model_h3_ols_rob,model_h3norm_ols_rob,
          show.se = T, #auto.label = T,
          collapse.se=T,
  col.order = c("est","se","p"),
  dv.labels = c("(1) Contribution Rate","(2) Activism Turnout", "(3.1) Welfare", "(3.2) Welfare/sec"),
  p.style = "numeric",digits = 2,
  p.adjust = "holm")


library(modelsummary)

# Save clean regression table for the paper (Table \ref{tab:mainreg})
save_reg_table(list("(1) Contribution Rate" = model_h1_ols_rob,
                     "(2) Activism Turnout"  = model_h2_main_rob,
                     "(3.1) Welfare"         = model_h3_ols_rob,
                     "(3.2) Welfare/sec"     = model_h3norm_ols_rob,
                     "Efficiency"            = model_efficiency_rob),
               path = paper_path("tables", "results", "mainreg.tex"))

```

#### Mixed Effects Models
```{r clean regression tables}

#main effects - OLS
tab_model(h1.lmer,h2.lmer,h3.lmer,h3b.lmer,
          show.se = T, #auto.label = T,
          collapse.se=T,
  col.order = c("est","se","p"),
  dv.labels = c("(1) Contribution Rate","(2) Activism Turnout", "(3.1) Welfare", "(3.2) Welfare/sec"),
  p.style = "numeric",digits = 2,
  p.adjust = "holm")


#main effects - mixed effects models
tab_model(h1.lmer,h2.lmer,h3.lmer,h3b.lmer,
          show.se = T, #auto.label = T,
          collapse.se=T,
  col.order = c("est","se","p"),
  dv.labels = c("(1) Contribution Rate","(2) Activism Turnout", "(3.1) Welfare", "(3.2) Welfare/sec"),
  p.style = "numeric",digits = 2,
  p.adjust = "holm")



#MAIN - interaction - mixed effects models - full sample
#H2 (Activism Turnout) uses h2.1_main.lmer: treatment_assigned + period only, matching
#the period-only specification used in mainreg.tex (lags tested separately in model_lag.tex)
tab_model(h1.1.lmer, h2.1_main.lmer,h3.1.lmer,h3.1b.lmer,
          show.se = T, #auto.label = T,
          collapse.se=T,
  col.order = c("est","se","p"),
  dv.labels = c("(1) Contribution Rate","(2) Activism Turnout", "(3.1) Welfare", "(3.2) Welfare/sec"),
  p.style = "numeric",digits = 3,
  p.adjust = "holm")

# Save clean regression table for the paper (Table A2, \ref{tab:lmer-interaction})
save_reg_table(list("(1) Contribution Rate" = h1.1.lmer,
                     "(2) Activism Turnout"  = h2.1_main.lmer,
                     "(3.1) Welfare"         = h3.1.lmer,
                     "(3.2) Welfare/sec"     = h3.1b.lmer,
                     "Efficiency"            = heff.lmer),
               path = paper_path("tables", "results", "lmer-interaction.tex"))


#interaction - mixed effects models - complete responses only
#H2 (Activism Turnout) uses h2.1clean_main.lmer: treatment_assigned + period only,
#matching the period-only specification used in the full-sample table
tab_model(h1.1clean.lmer, h2.1clean_main.lmer,h3.1clean.lmer,h3.1bclean.lmer,
          show.se = T, #auto.label = T,
          collapse.se=T,
  col.order = c("est","se","p"),
  dv.labels = c("(1) Contribution Rate","(2) Activism Turnout", "(3.1) Welfare", "(3.2) Welfare/sec"),
  p.style = "numeric",digits = 3,
  p.adjust = "holm")

# Save clean regression table for the paper (Table A5, \ref{tab:mainreg_controls})
save_reg_table(list("(1) Contribution Rate" = h1.1clean.lmer,
                     "(2) Activism Turnout"  = h2.1clean_main.lmer,
                     "(3.1) Welfare"         = h3.1clean.lmer,
                     "(3.2) Welfare/sec"     = h3.1bclean.lmer,
                     "Efficiency"            = heffclean.lmer),
               path = paper_path("tables", "results", "lmer-interaction-comp.tex"))

```

# Exploratory Analyses

## Robustness: Controls (Age, Gender, Education, etc.)

Moved to its own script, `03_robustness_controls.Rmd` -- builds Tables A3
(`tables/results/mainreg_controls2.tex`) and A4
(`tables/results/mainreg_more_controls2.tex`) from `df_long`'s own
questionnaire columns.

## Temporal Dynamics: Contribution Trends
```{r temporal contrib}

#contribution rate over time - interaction terms not sig (no sig differences in contribution rates over time across conditions (but general decline over time (ME)))
model_trend <- lm(relative_contribution ~ condition * period,
                  data = df_long)
vcov_trend <- vcovCL(model_trend, cluster = ~ group_id)
coeftest(model_trend, vcov = vcov_trend)


interact_plot(model_trend,modx = condition, pred = period,
              #plot.points = T,
              rug = T,jitter = .1,
              #interval = T,
              robust = T,line.thickness = 2,int.type = "confidence",
              colors = condition_colors) + 
  theme_classic()

```

```{r temporal activism}

# Activism turnout over time (treatment rounds only) - no sig difference over time
model_trend_activism <- lm(
  activism_rate ~ condition * period,
  data = df_long %>% filter(is_treatment_round == 1, !is.na(chose_activism)))
vcov_trend_activism <- vcovCL(model_trend_activism, cluster = ~ group_id)
coeftest(model_trend_activism, vcov = vcov_trend_activism)

```

```{r to do, eval=FALSE, include=FALSE}

#Table x. Group Activism Norms ~ Activism Participation in Subsequent Round
h.norms.lmer <- lmer(lag(num_activists) ~ chose_activism + treatment_assigned + period + (1 | group_id) + (1 | participant.code), 
           data = df_long %>% filter(is_treatment_round == 1))
tab_model(h.norms.lmer)

df_long$num_activists


#Table x. Individual Participation ~ Contribution Rate
h.pcont.lmer <- lmer(chose_activism ~ relative_contribution + treatment_assigned + period + (1 | group_id) + (1 | participant.code), 
           data = df_long %>% filter(is_treatment_round == 1))
tab_model(h.pcont.lmer)


#Table x. Wage ~ Contribution
h.wagecont.lmer <- lmer(wage ~ relative_contribution + treatment_assigned + period + (1 | group_id) + (1 | participant.code), 
           data = df_long %>% filter(is_treatment_round == 1))
tab_model(h.wagecont.lmer)


#Table x. Who participates in activism? Previous round performance + contributions
h.pcont.lmer <- lmer(chose_activism ~ relative_contribution + treatment_assigned + period + (1 | group_id) + (1 | participant.code), 
           data = df_long %>% filter(is_treatment_round == 1))
tab_model(h.pcont.lmer)

```

```{r power, eval=FALSE, include=FALSE}
library(simr)
library(performance)

# Primary analysis uses df_long_clean (excludes 5 groups with technical failures)
# df_long retained for robustness comparison only
#takes a few mins to run so run manually

# --- Fit model on complete sessions only ---
h1.clean <- lmer(
  relative_contribution ~ treatment_assigned * is_treatment_round + period +
    (1 | group_id) + (1 | participant.code),
  data = df_long_clean
)
summary(h1.clean)

# --- Variance components & ICC ---
cat("=== df_long_clean (primary) ===\n")
performance::icc(h1.clean)
VarCorr(h1.clean)

cat("\n=== df_long (robustness check) ===\n")
performance::icc(h1.1.lmer)

# --- Conservative power simulation ---
# Halve all interaction effects to correct for small-sample inflation
# Demonstration is reference level; its treatment-round effect = is_treatment_round coef
# Unit of randomisation: groups (4 participants/group, 14 rounds each)
set.seed(42)

model_power <- extend(h1.clean, along = "group_id", n = 120)

fixef(model_power)["treatment_assignedpetition:is_treatment_round"]  <- fixef(h1.clean)["treatment_assignedpetition:is_treatment_round"] / 2
fixef(model_power)["treatment_assignedroadblock:is_treatment_round"] <- fixef(h1.clean)["treatment_assignedroadblock:is_treatment_round"] / 2
fixef(model_power)["is_treatment_round"]                             <- fixef(h1.clean)["is_treatment_round"] / 2

cat("\nConservative effect sizes (halved from pilot, complete sessions):\n")
cat("  Demonstration in treatment rounds:", round(fixef(model_power)["is_treatment_round"], 2), "\n")
cat("  Petition vs Demo (interaction):   ", round(fixef(model_power)["treatment_assignedpetition:is_treatment_round"], 2), "\n")
cat("  Roadblock vs Demo (interaction):  ", round(fixef(model_power)["treatment_assignedroadblock:is_treatment_round"], 2), "\n\n")

# --- Power curve: 12–40 groups ---
pc_h1 <- powerCurve(
  model_power,
  test     = fixed("treatment_assigned:is_treatment_round", "lr"),
  along    = "group_id",
  breaks   = seq(12, 40, by = 4),
  nsim     = 500,
  progress = FALSE
)

pc_power  <- sapply(pc_h1$ps, function(x) mean(x$pval < x$alpha, na.rm = TRUE))
pc_groups <- pc_h1$nlevels
nice_table(data.frame(groups = pc_groups, participants = pc_groups * 4, power = round(pc_power, 3)))

plot(pc_h1)

# --- Sample size recommendation ---
above80 <- pc_groups[pc_power >= 0.80]
above95 <- pc_groups[pc_power >= 0.95]
mg80    <- if (length(above80) > 0) min(above80) else NA
mg95    <- if (length(above95) > 0) min(above95) else NA

cat("\n--- Sample Size Recommendation (df_long_clean, halved effects) ---\n")
cat("ICC (adjusted):              ", round(performance::icc(h1.clean)$ICC_adjusted, 3), "\n")
cat("Design effect (group size=4):", round(1 + 3 * performance::icc(h1.clean)$ICC_adjusted, 2), "\n\n")
cat("80% power: ", mg80, "groups /", mg80 * 4, "participants /", ceiling(mg80 * 4 / 3), "per condition\n")
cat("95% power: ", mg95, "groups /", mg95 * 4, "participants /", ceiling(mg95 * 4 / 3), "per condition\n")
cat("95% + 20% attrition buffer:", ceiling(mg95 * 1.2), "groups /",
    ceiling(mg95 * 1.2) * 4, "participants /",
    ceiling(ceiling(mg95 * 1.2) * 4 / 3), "per condition\n")

```

```{r git-commit, eval=FALSE, include=FALSE}

repo_root <- trimws(system("git rev-parse --show-toplevel", intern = TRUE))
git <- function(...) system2("git", c("-C", shQuote(repo_root), shQuote(c(...))))

git("add", "analysis/", "tables/", "figures/")
git("commit", "-m", paste0("Update analysis output - ", Sys.time()))

# Pull in any remote changes (e.g. files added via github.com) before pushing,
# so a diverged remote fails loudly here instead of silently dropping the push.
git("fetch", "origin", "main")
behind <- as.integer(trimws(system2("git",
  c("-C", shQuote(repo_root), "rev-list", "--count", "HEAD..origin/main"), stdout = TRUE)))

if (behind > 0) {
  merge_status <- git("merge", "origin/main", "--no-edit")
  if (merge_status != 0) {
    stop("git-commit chunk: merge with origin/main failed, likely a conflict. ",
         "Resolve manually (git status) in a terminal, then re-run this chunk.")
  }
}

push_status <- git("push", "origin", "main")
if (push_status != 0) {
  stop("git-commit chunk: push to origin/main failed. ",
       "Run `git status` / `git log --oneline -5` manually to diagnose.")
}

```
