---
title: "Activism Pilot Analysis -- Robustness: Sociodemographic Controls"
author: "Shaye"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
    theme: flatly
---

# Set Up

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
options(scipen = 999, digits = 4)

suppressPackageStartupMessages({
  library(tidyverse)
  library(here)
  library(estimatr)
  library(modelsummary)
})

# Robust path resolver across VS Code, RStudio, and command-line execution
find_path <- function(...) {
  p1 <- here::here(...)
  if (file.exists(p1) || dir.exists(p1)) return(p1)

  p2 <- here::here("analysis", ...)
  if (file.exists(p2) || dir.exists(p2)) return(p2)

  p3 <- file.path(...)
  if (file.exists(p3) || dir.exists(p3)) return(p3)

  p4 <- file.path("..", ...)
  if (file.exists(p4) || dir.exists(p4)) return(p4)

  return(p1)
}

# Repo root -- tables/results is git-tracked at the repo root and read by
# paper.tex via root-relative paths, regardless of where here::here() resolves
repo_root <- trimws(system("git rev-parse --show-toplevel", intern = TRUE))
paper_path <- function(...) file.path(repo_root, ...)

# Holm-adjusted p-values for every lm_robust model passed to modelsummary(),
# applied per model across all non-intercept terms -- matches
# 02_pilot_analysis.Rmd's convention for all other regression tables in the paper.
tidy_custom.lm_robust <- function(x, ...) {
  pp <- as.data.frame(parameters::model_parameters(x))
  is_intercept <- pp$Parameter == "(Intercept)"
  p_adj <- pp$p
  p_adj[!is_intercept] <- p.adjust(p_adj[!is_intercept], method = "holm")
  data.frame(term = pp$Parameter, p.value = p_adj)
}

# Pretty term labels matching the paper's established style
REG_TERM_LABELS <- c(
  "conditionpetition"       = "condition [petition]",
  "conditiondemonstration"  = "condition [demonstration]",
  "conditionroadblock"      = "condition [roadblock]",
  "conditioncontrol"        = "condition [control]",
  "female"                                 = "Female",
  "Education_Years"                        = "Education (years)",
  "Monthly_Disposable_Income_EUR_numeric"  = "Monthly income (EUR)",
  "donated"                                = "Donated",
  "Political_Orientation_0_10"             = "Political orientation (0-10)",
  "General_Trust_0_10"                     = "General trust (0-10)",
  "Risk_Willingness_0_10"                  = "Risk willingness (0-10)"
)

# modelsummary's LaTeX output wraps the table in its own \begin{table}...\end{table};
# paper.tex already wraps each \input{} in its own table float (caption, label,
# resizebox), so strip that outer wrapper here to avoid nesting two table floats.
strip_table_wrapper <- function(path) {
  ln <- readLines(path)
  ln <- ln[!ln %in% c("\\begin{table}", "\\centering", "\\end{table}")]
  writeLines(ln, path)
}

# estimate/statistic/shape put p in its own column next to "Estimate (SE)", matching
# the paper's established table style, instead of modelsummary's default of stacking
# SE and p together underneath the estimate within a single per-model column.
save_reg_table <- function(models, path, coef_rename = REG_TERM_LABELS,
                            estimate = "{estimate} ({std.error})",
                            statistic = "p.value",
                            shape = term ~ model + statistic,
                            ...) {
  modelsummary::modelsummary(models, output = path, coef_rename = coef_rename,
                              estimate = estimate, statistic = statistic, shape = shape, ...)
  strip_table_wrapper(path)
}
```

# Load Processed Data

```{r load data}
# Load processed clean datasets (generated by 01_data_cleaning.Rmd)
load(find_path("data", "processed", "clean_dfs.RData"))
```

# Robustness: Sociodemographic Controls

Adds demographic controls (gender, education, income, donation history, and --
in the extended specification -- political orientation, general trust, and risk
willingness) to the main H1/H2/H3 specifications, as a robustness check on the
main results in `02_pilot_analysis.Rmd`. Uses the full sample (`df_long`),
matching `tables/results/mainreg.tex`.

```{r robustness controls}

# Demographic controls are collected in df_long's own questionnaire columns
# (no separate df_questionnaire join needed). Education_Years and donated are
# derived categorical->ordinal/binary recodes; income parses "low-high" range
# strings to their midpoint.
df_long_demo <- df_long %>%
  mutate(
    female = case_when(
      Gender == "Woman" ~ 1,
      Gender == "Man" ~ 0,
      TRUE ~ NA_real_
    ),
    Education_Years = case_when(
      Year_of_Study_Semester == "1-2" ~ 1,
      Year_of_Study_Semester == "3-4" ~ 2,
      Year_of_Study_Semester == "5-6" ~ 3,
      Year_of_Study_Semester %in% c("7-8", "9+") ~ 4,
      TRUE ~ NA_real_
    ),
    Monthly_Disposable_Income_EUR_numeric = sapply(Monthly_Disposable_Income_EUR, function(x) {
      if (is.na(x)) return(NA_real_)
      if (str_detect(x, "-")) return(mean(as.numeric(str_split(x, "-", simplify = TRUE))))
      as.numeric(x)
    }),
    donated = case_when(
      Charitable_Donations == "No" ~ 0,
      Charitable_Donations %in% c("Once", "Regularly") ~ 1,
      TRUE ~ NA_real_
    )
  )

# --- Table A3: sociodemographic controls (Table \ref{tab:mainreg_controls}) ---
model_h1_ctrl2 <- estimatr::lm_robust(relative_contribution ~ condition + female + Education_Years + Monthly_Disposable_Income_EUR_numeric + donated + period,
                   data = df_long_demo,
                   clusters = group_id,
                   se_type = "CR2")

# Reference = demonstration for Activism Turnout: control has zero valid rows
# for this outcome, so a control reference leaves the model rank-deficient
# (see 02_pilot_analysis.Rmd's model_h2_ols_rob for the full diagnosis).
model_h2_ctrl2 <- estimatr::lm_robust(player.chose_activism ~ condition + female + Education_Years + Monthly_Disposable_Income_EUR_numeric + donated + period,
                   data = df_long_demo %>% mutate(condition = relevel(factor(condition), ref = "demonstration")),
                   clusters = group_id,
                   se_type = "CR2")

model_h3_ctrl2 <- estimatr::lm_robust(welfare ~ condition + female + Education_Years + Monthly_Disposable_Income_EUR_numeric + donated + period,
                   data = df_long_demo,
                   clusters = group_id,
                   se_type = "CR2")

model_h3norm_ctrl2 <- estimatr::lm_robust(norm_welfare ~ condition + female + Education_Years + Monthly_Disposable_Income_EUR_numeric + donated + period,
                   data = df_long_demo,
                   clusters = group_id,
                   se_type = "CR2")

save_reg_table(list("(1) Contribution Rate" = model_h1_ctrl2,
                     "(2) Activism Turnout"  = model_h2_ctrl2,
                     "(3.1) Welfare"         = model_h3_ctrl2,
                     "(3.2) Welfare/sec"     = model_h3norm_ctrl2),
               path = paper_path("tables", "results", "mainreg_controls2.tex"))

# --- Table A4: + political orientation, trust, risk willingness (Table \ref{tab:mainreg_more_controls}) ---
model_h1_morectrl2 <- estimatr::lm_robust(relative_contribution ~ condition + female + Education_Years + Monthly_Disposable_Income_EUR_numeric + donated + Political_Orientation_0_10 + General_Trust_0_10 + Risk_Willingness_0_10 + period,
                   data = df_long_demo,
                   clusters = group_id,
                   se_type = "CR2")

model_h2_morectrl2 <- estimatr::lm_robust(player.chose_activism ~ condition + female + Education_Years + Monthly_Disposable_Income_EUR_numeric + donated + Political_Orientation_0_10 + General_Trust_0_10 + Risk_Willingness_0_10 + period,
                   data = df_long_demo %>% mutate(condition = relevel(factor(condition), ref = "demonstration")),
                   clusters = group_id,
                   se_type = "CR2")

model_h3_morectrl2 <- estimatr::lm_robust(welfare ~ condition + female + Education_Years + Monthly_Disposable_Income_EUR_numeric + donated + Political_Orientation_0_10 + General_Trust_0_10 + Risk_Willingness_0_10 + period,
                   data = df_long_demo,
                   clusters = group_id,
                   se_type = "CR2")

model_h3norm_morectrl2 <- estimatr::lm_robust(norm_welfare ~ condition + female + Education_Years + Monthly_Disposable_Income_EUR_numeric + donated + Political_Orientation_0_10 + General_Trust_0_10 + Risk_Willingness_0_10 + period,
                   data = df_long_demo,
                   clusters = group_id,
                   se_type = "CR2")

save_reg_table(list("(1) Contribution Rate" = model_h1_morectrl2,
                     "(2) Activism Turnout"  = model_h2_morectrl2,
                     "(3.1) Welfare"         = model_h3_morectrl2,
                     "(3.2) Welfare/sec"     = model_h3norm_morectrl2),
               path = paper_path("tables", "results", "mainreg_more_controls2.tex"))

```
